### 2018-03-17 paper/7-related.tex
Tweak memoization

As a simple example of using a build system for memoization, we solve a textbook
dynamic programming problem~--~Levenshtein's \emph{edit
distance}~\cite{levenshtein1966binary}: given two input strings $a$ and
$b$, find the minimum-weight series of edit operations that transforms $a$
to $b$. The edit operations are typically \emph{inserting}, \emph{deleting} or
\emph{replacing} a symbol. The dynamic programming solution of this problem is
widely known (e.g., see~\cite{cormen2001introduction}) that we provide its
encoding in our \hs{Task} abstraction without further explanation.
We address elements of strings $a_i$ and $b_i$ by keys \hs{A}~$i$ and \hs{B}~$i$,
respectively, while the cost of a subproblem $d_{ij}$ is identified by
\hs{D}~$i$~$j$.

\begin{minted}[xleftmargin=10pt]{haskell}
data Key = A Integer | B Integer | D Integer Integer deriving Ord
\end{minted}
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
editDistance :: Task Monad Key Integer
editDistance _     (D i 0) = Just $ pure i
editDistance _     (D 0 j) = Just $ pure j
editDistance fetch (D i j) = Just $ do
    ai <- fetch (A i)
    bj <- fetch (B i)
    if ai == bj
        then fetch (D (i - 1) (j - 1))
        else do
            x <- fetch (D (i - 1)  j     )
            y <- fetch (D  i      (j - 1))
            z <- fetch (D (i - 1) (j - 1))
            return (1 + minimum [x, y, z])
editDistance _ _ = Nothing
\end{minted}

\noindent
When asked to build \hs{D}~$n$~$m$, a minimal build system will calculate the
result using memoization. Furthermore, when an input symbol $a_i$ is changed,
only necessary, incremental recomputation will be performed.

>>>>

As a simple example of using a build system for memoization, we solve a textbook
dynamic programming problem~--~Levenshtein's \emph{edit
distance}~\cite{levenshtein1966binary}: given two input strings $a$ and
$b$, find the minimum-weight series of edit operations that transforms $a$
to $b$. The edit operations are typically \emph{inserting}, \emph{deleting} or
\emph{replacing} a symbol. The dynamic programming solution of this problem is
so widely known (e.g., see~\cite{cormen2001introduction}) that we provide its
encoding in our \hs{Task} abstraction without further explanation.
We address elements of strings $a_i$ and $b_i$ by keys \hs{A}~$i$ and \hs{B}~$i$,
respectively, while the cost of a subproblem $d_{ij}$ is identified by
\hs{D}~$i$~$j$.

\begin{minted}[xleftmargin=10pt]{haskell}
data Key = A Integer | B Integer | D Integer Integer deriving Ord
\end{minted}
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
editDistance :: Task Monad Key Integer
editDistance _     (D i 0) = Just $ pure i
editDistance _     (D 0 j) = Just $ pure j
editDistance fetch (D i j) = Just $ do
    ai <- fetch (A i)
    bj <- fetch (B i)
    if ai == bj
        then fetch (D (i - 1) (j - 1))
        else do
            insert  <- fetch (D  i      (j - 1))
            delete  <- fetch (D (i - 1)  j     )
            replace <- fetch (D (i - 1) (j - 1))
            return (1 + minimum [insert, delete, replace])
editDistance _ _ = Nothing
\end{minted}

\noindent
When asked to build \hs{D}~$n$~$m$, a minimal build system will calculate the
result using memoization. Furthermore, when an input symbol $a_i$ is changed,
only necessary, incremental recomputation will be performed~--~an optimisation
that cannot be achieved just with memoization.
